---
title: "Goodplot and Badplot"
author: "Leah"
date: "2023-03-28"
output: 
  prettydoc::html_pretty:
    theme: hpstr
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      fig.path = "../output/") 
```

# Introduction 
Today we will be looking at my goodplot and badplot that I worked really hard on. 

# Load Libraries 
```{r}
library(ggplot2)
library(readr)
library(dplyr)
library(here)
library(prettydoc)
```

# Load Data 
```{r}
us_births_2000_2014 <- read_csv(here("Goodplot_and_Badplot", "Data", "us_births_2000-2014.csv"))
View(us_births_2000_2014)
```
# **What Makes This a Bad Plot**
- Too many points
- Unclear legends 
- Missing axis titles 
- Misleading axes ranges
- Overuse of colors  
- Not clear it's cluttered 

#BadPlot
```{r Badplot Updated}
us_births_2000_2014 %>% #plot data 
  ggplot(aes(x = births, y = year))+
  geom_point(
    color = "blue", # point color, shape, and size 
    fill = "orange",
    shape = 23,
    alpha = 0.5,
    size = 5,
    stroke = 2
  )+
  facet_wrap(~month) + #look at it by month 
  labs( x = "", y = "") + # remove axis titles 
  scale_y_continuous(breaks = c(2005)) + # only label 2005 
  theme_bw() +
  theme(panel.background = element_rect(fill = "purple")) # theme and background 
```

# **What Makes A Goodplot**
- Axes are clean and clear
- Caption to explain plot
- Emphasis on births in the year 
- Appropriate graph 
- Clear/ not cluttered 
- Nice color scheme 


#Goodplot 
```{r Yearly Births vs. Month, fig.height = 8, fig.width = 12, fig.cap = "US births for the years 2000 to 2014."}
us_births_2000_2014 %>%
ggplot(aes(x = month, y = births)) +
  geom_bar(stat = "identity", fill = "blue", alpha = 1, width =.4) +
  coord_flip () + # flip coordinate 
  facet_wrap(~year) + # look at by year 
  labs(x = "Month", y = "Births", title = "Yearly Births vs. Month") +
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12), 
                     labels = c("Jan", "Feb", "Mar","Apr",  "May","Jun","Jul","Aug","Sep","Oct", "Nov", "Dec")) + # number tto word 
  theme_dark() +
  theme(panel.background = element_rect(fill = "snow2")) + # panel background 
theme(plot.background = element_rect(fill = "lightblue")) + # background 
  theme(axis.title = element_text(size = 20,
                                  color = "navy"))
```








