---
title: "Tuesday Class"
author: "Leah"
date: "2023-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries 

```{r}
library(tidyverse)
library(here)
```

#Load Data
```{r}
testdata<-read_csv(here("Week_12", "Data","cond data","011521_CT316_1pcal.csv"))
glimpse(testdata)
```


# Simple for loop
```{r}
print(paste("The year is", 2000))

years<-c(2015:2021)
for (i in years){ # set up the for loop where i is the index
  print(paste("The year is", i)) # loop over i
}

#Pre-allocate space for the for loop
# empty matrix
year_data<-data.frame(matrix(ncol = 2, nrow = length(years)))
# add column names
colnames(year_data)<-c("year", "year_name")
year_data

for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over i
}
year_data

for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over year name
  year_data$year[i]<-years[i] # loop over year
}
year_data

```

# List files in a directory 
```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_12", "Data", "cond data")

files <- dir(path = CondPath,pattern = ".csv")
files
```

# Pre-allocate space for the loop 
```{r}
# pre-allocate space
# make an empty dataframe that has one row for each file and 3 columns
cond_data<-data.frame(matrix(nrow = length(files), ncol = 3))
# give the dataframe column names
colnames(cond_data)<-c("filename","mean_temp", "mean_sal")
cond_data
```

# For loop 
```{r}
raw_data<-read_csv(paste0(CondPath,"/",files[1])) # test by reading in the first file and see if it works
head(raw_data)

mean_temp<-mean(raw_data$Temperature, na.rm = TRUE) # calculate a mean
mean_temp
```
# Turn it into a for loop

```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files
}
```

# Add in the loop over the raw data 

```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
glimpse(raw_data)
}
```
# Add in the columns 
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
} 
cond_data
```
# Add in means 
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
cond_data$mean_temp[i]<-mean(raw_data$Temperature, na.rm =TRUE)
cond_data$mean_sal[i]<-mean(raw_data$Salinity, na.rm =TRUE)
} 
cond_data
```
# Simple example of map function 
```{r}
1:10 %>% # a vector from 1 to 10 (we are going to do this 10 times)
  map(rnorm, n = 15) %>% # calculate 15 random numbers based on a normal distribution in a list
  map_dbl(mean) # calculate the mean. It is now a vector which is type "double"
  
```

# Make your own function 
```{r}
1:10 %>% # list 1:10
  map(function(x) rnorm(15, x)) %>% # make your own function
  map_dbl(mean)
```

# Use a formula when you want to chnage the arguments within the function 

```{r}
1:10 %>%
  map(~ rnorm(15, .x)) %>% # changes the arguments inside the function
  map_dbl(mean)
```

# Bring in files using purrr instead of a for loop 
```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_12", "Data", "cond data")
files <- dir(path = CondPath,pattern = ".csv")
files

# Or we can get the full file names in one less step 

files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files

```
# Read in the files 
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename
data
```
# Calculate means 
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temperature, na.rm = TRUE),
            mean_sal = mean(Salinity,na.rm = TRUE))
data
```

