---
title: "Thursday Class"
author: "Leah"
date: "2023-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Today We will be working with factors #

# Load Libraries 
```{r}
library(tidyverse)
library(here)
```

# Load Data 

```{r}
income_mean <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_mean.csv')
```
# To make something a factor you put the vector in the function factor()
```{r}
fruits<-factor(c("Apple", "Grape", "Banana"))
fruits
```

# Booby trap 
```{r}
test<-c("A", "1", "2")
as.numeric(test)
```

# Test to see if it was a factor 
```{r}
test<-factor(test) # covert to factor
as.numeric(test)
```

# Starwars 
```{r}
glimpse(starwars)

starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) # how many individuals of each species are present across all starwars films?


```
# Lump all the species together that had less than 3 individuals
```{r}
star_counts<-starwars %>%
  filter(!is.na(species)) %>%
  mutate(species = fct_lump(species, n = 3)) %>%
  count(species)
star_counts
```

# Reordering factors 
```{r}
star_counts %>%
  ggplot(aes(x = species, y = n))+
  geom_col()
```
# How would you view them in order from lowest to highest?
```{r}
star_counts %>%
  ggplot(aes(x = fct_reorder(species, n), y = n))+ # reorder the factor of species by n
  geom_col()
```
# How would you make it descending?
```{r}
star_counts %>%
  ggplot(aes(x = fct_reorder(species, n, .desc = TRUE), y = n))+ # reorder the factor of species by n
  geom_col() +
  labs(x = "Species")
```
# Reorder line plots 
```{r}
glimpse(income_mean)

total_income<-income_mean %>%
  group_by(year, income_quintile)%>%
  summarise(income_dollars_sum = sum(income_dollars))%>%
  mutate(income_quintile = factor(income_quintile)) # make it a factor


total_income%>%
ggplot(aes(x = year, y = income_dollars_sum, color = income_quintile))+
  geom_line()

total_income%>%
  ggplot(aes(x = year, y = income_dollars_sum, 
             color = fct_reorder2(income_quintile,year,income_dollars_sum)))+
  geom_line()+
  labs(color = "income quantile")
```
# Let's say we have a vector and we want to put it in a specific order that we say.
```{r}
x1 <- factor(c("Jan", "Mar", "Apr", "Dec"))
x1

x1 <- factor(c("Jan", "Mar", "Apr", "Dec"), levels = c("Jan", "Mar", "Apr", "Dec"))
x1
```
# Let's say we instead of grouping our species that have < 3 counts into "Other" we just filter them out.
```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor
  filter(n>3) # only keep species that have more than 3
starwars_clean

levels(starwars_clean$species)
```
# Drop levels 
```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor 
  filter(n>3)  %>% # only keep species that have more than 3 
  droplevels() # drop extra levels
levels(starwars_clean$species)
```

# If you want to rename (or recode) a level
```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor 
  filter(n>3)  %>% # only keep species that have more than 3 
  droplevels() %>% # drop extra levels 
  mutate(species = fct_recode(species, "Humanoid" = "Human"))
starwars_clean
```

